---
title: "p8105_hw2_wq2161"
author: "Wanxin Qi"
date: "10/4/2021"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
```

## Problem 1

### Part 1

Read and clean the Mr. Trash Wheel sheet.

```{r trash_wheel_df}
trash_wheel_df = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N535")

trash_wheel_df =
  trash_wheel_df %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = round(sports_balls))

trash_wheel_df
```

### Part 2

Read and clean precipitation data for 2018 and 2019.

```{r precipitation_datasets}
prcp_2018 = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2018 Precipitation", range = "A2:B15")

prcp_2018 =
  prcp_2018 %>%
  janitor::clean_names() %>%
  drop_na(month, total) %>%
  mutate(year = "2018") %>%
  relocate(year)

prcp_2019 = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2019 Precipitation", range = "A2:B15")

prcp_2019 =
  prcp_2019 %>%
  janitor::clean_names() %>%
  drop_na(month, total) %>%
  mutate(year = "2019") %>%
  relocate(year)
```

Combine precipitation datasets and convert month to a character variable.

```{r prcp_2018_2019}
prcp_2018_2019 = 
  bind_rows(prcp_2018, prcp_2019) %>%
  mutate(month = month.name[month])

prcp_2018_2019
```

### Part 3

The datasets are collected by *Mr. Trash Wheel*. They are under the background of “a water-wheel vessel that removes trash from the Inner Harbor in Baltimore, Maryland.” The trash_wheel_df dataset contains `r ncol(trash_wheel_df)` variables and `r nrow(trash_wheel_df)` observations. It includes information on `r colnames(trash_wheel_df)`. The prcp_2018_2019 dataset is the combination of the total precipitation amounts by month in 2018 and 2019. The variables are `r colnames(prcp_2018_2019)`. It contains `r nrow(prcp_2018_2019)` observations. The total precipitation in 2018 was `r prcp_2018_2019 %>% filter(year == "2018") %>% pull(total) %>% sum()`. The median number of sports balls in a dumpster in 2019 is `r trash_wheel_df %>% filter(year == c("2019")) %>% pull(sports_balls) %>% median()`.


## Problem 2

### Part 1

Read and clean the data in pols-month.csv.

```{r pols_month_df, message = FALSE}
pols_month_df = read_csv("data/fivethirtyeight_datasets/pols-month.csv")

pols_month_df =
  pols_month_df %>%
  janitor::clean_names() %>%
  mutate(mon = ymd(mon)) %>%
  separate(col = mon, into = c("year", "month", "day"), sep = "-") %>%
  mutate(
    year = as.numeric(year),
    month = as.numeric(month),
    month = month.name[month],
    president = ifelse(prez_dem == "1", "dem", "gop")
  ) %>%
  select(-prez_dem, -prez_gop, -day) %>%
  relocate(year, month, president)

pols_month_df
```

### Part 2

Read and clean the data in snp.csv.

```{r snp_df, message = FALSE}
snp_df = read_csv("data/fivethirtyeight_datasets/snp.csv")

snp_df =
  snp_df %>%
  janitor::clean_names() %>%
  mutate(date = mdy(date)) %>%
  separate(col = date, into = c("year", "month", "day"), sep = "-") %>%
  mutate(
    year = as.numeric(year),
    year = ifelse(year > 2021, year - 100, year)
  ) %>%
  arrange(year, month) %>%
  mutate(
    year = as.numeric(year),
    month = as.numeric(month),
    month = month.name[month]
  ) %>%
  select(-day) %>%
  relocate(year, month)

snp_df
```

### Part 3

Tidy the data in unemployment.csv.

```{r unemployment_df, message = FALSE}
unemployment_df = read_csv("data/fivethirtyeight_datasets/unemployment.csv")

unemployment_df =
  unemployment_df %>%
  janitor::clean_names()

unemployment_tidy =
  unemployment_df %>%
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "percentage"
  ) %>%
  drop_na() %>%
  mutate(
    month = replace(month, month == "jan", "January"),
    month = replace(month, month == "feb", "February"),
    month = replace(month, month == "mar", "March"),
    month = replace(month, month == "apr", "April"),
    month = replace(month, month == "may", "May"),
    month = replace(month, month == "jun", "June"),
    month = replace(month, month == "jul", "July"),
    month = replace(month, month == "aug", "August"),
    month = replace(month, month == "sep", "September"),
    month = replace(month, month == "oct", "October"),
    month = replace(month, month == "nov", "November"),
    month = replace(month, month == "dec", "December")
  ) %>%
  relocate(year, month)

unemployment_tidy
```

### Part 4

Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r join_datasets}
pols_month_date_df =
  pols_month_df %>%
  mutate(with_day = as.Date(paste(month, "01", year, sep = "-"), format = "%b-%d-%Y")) %>%
  mutate(date = format(as.Date(with_day), "%Y-%m")) %>%
  select(-with_day) %>%
  relocate(date)

snp_date_df =
  snp_df %>%
  mutate(with_day = as.Date(paste(month, "01", year, sep = "-"), format = "%b-%d-%Y")) %>%
  mutate(date = format(as.Date(with_day), "%Y-%m")) %>%
  select(-with_day) %>%
  relocate(date)

unemployment_date_tidy =
  unemployment_tidy %>%
  mutate(with_day = as.Date(paste(month, "01", year, sep = "-"), format = "%b-%d-%Y")) %>%
  mutate(date = format(as.Date(with_day), "%Y-%m")) %>%
  select(-with_day) %>%
  relocate(date)

pols_snp_df =
  left_join(pols_month_date_df, snp_date_df, by = c("date" = "date")) %>%
  select(-year.x, -month.x, -year.y, -month.y)

pols_snp_unemployment_df =
  left_join(pols_snp_df, unemployment_date_tidy, by = c("date" = "date")) %>%
  select(-year, -month)

pols_snp_unemployment_df
```

### Part 5

The pols-month dataset contains `r ncol(pols_month_df)` variables and `r nrow(pols_month_df)` observations related to the number of national politicians who are democratic or republican at any given time. The range of years starts from `r min(pull(pols_month_df, year))` to `r max(pull(pols_month_df, year))`. It includes information of `r colnames(pols_month_df)`.

The snp dataset contains `r ncol(snp_df)` variables and `r nrow(snp_df)` observations related to Standard & Poor’s stock market index (S&P), often used as a representative measure of stock market as a whole. The range of years starts from `r min(pull(snp_df, year))` to `r max(pull(snp_df, year))`. It includes information of `r colnames(snp_df)`.

The unemployment dataset contains `r ncol(unemployment_tidy)` variables and `r nrow(unemployment_tidy)` observations related to percentage of unemployment. The range of years starts from `r min(pull(unemployment_tidy, year))` to `r max(pull(unemployment_tidy, year))`. It includes information of `r colnames(unemployment_tidy)`.

The pols_snp_unemployment dataset is a combination of the three datasets, which contains `r ncol(pols_snp_unemployment_df)` variables and `r nrow(pols_snp_unemployment_df)` observations. The range of years starts from `r min(substr(pull(pols_snp_unemployment_df, date), 1, 4))` to `r max(substr(pull(pols_snp_unemployment_df, date), 1, 4))`. It includes information of `r colnames(pols_snp_unemployment_df)`.


